---
title: Credit transfer service benchmarking
date: 2021-06-08
---

```{r report-config, include = F}
opts_chunk$set(error = F, dev = "svg")
```

# Executive summary

1. The most performant and the most cost effective instance type is c6g
1. Performance linearly increases with the number of instances
1. Cost efficiency increases exponentially with the number of instances

# Introduction

The credit transfer service (CTS) benchmarking analyses the measurements and results
registred during the benchmarking. Both **single node** and **multi-node** CTS
configurations are analysed.

The **cost analysis** consists in comparing the cost of 1M transactions (TX) of a given
instance type or a configured number of instances. The **performance analysis**
comprises the comparison of number of transactions per second (TPS) for each instance
type or an employed number of instances. The **cost efficiency analysis** compares the
performance of the CTS agains the cost of 1M TX of a specific instance type of the
number of used instances.

# Transaction cost and performance analysis

## Single node pricing

```{r snp-data, include = F}
snp <- data.frame(
  inst.type = c("c5", "c5a", "c5n", "c6g"),
  tps.peak = c(505, 469, 509, 572),
  cost.hour = c(0.77, 0.69, 0.98, 0.61),
  cost.mtx = c(0.42, 0.41, 0.53, 0.30))
```

```{r snp-data-table, echo = F}
kable(snp, caption = "Single node pricing and TPS data",
  col.names = c("Instance type", "TPS peak", "Hour cost, $", "1M TX cost, $"),
  align = "crrr")
```

### Performance of instance type


```{r snp-tps.peak-inst.type, fig.cap = "TPS peak per instance type (single node)"}
ggplot(snp, aes(inst.type, tps.peak,
  color = inst.type, fill = inst.type)) +
  geom_bar(stat = "identity", alpha = 0.2) +
  xlab("Instance type") + ylab("TPS peak")
```

### Cost of 1M transactions of instance type


```{r snp-cost.mtx-inst.type, fig.cap = "Cost of 1M TX per instance type (single node)"}
ggplot(snp, aes(inst.type, cost.mtx,
  color = inst.type, fill = inst.type)) +
  geom_bar(stat = "identity", alpha = 0.2) +
  xlab("Instance type") + ylab("Cost of 1M TX")
```

### Cost efficiency of instance type


```{r snp-cost.mtx-tps.peak, fig.cap = "TPS peak vs cost of 1M TX per instance type (single node)"}
ggplot(snp, aes(cost.mtx, tps.peak)) +
  geom_point(aes(size = tps.peak, color = tps.peak)) +
  scale_size_continuous(range = c(5, 12)) +
  geom_text(aes(x = cost.mtx, y = tps.peak, label = inst.type),
    hjust = "left", nudge_x = 0.015, size = 5) +
  xlim(0.28, 0.6) + xlab("Cost of 1M TX") + ylab("TPS peak")
```

## Multi-node pricing


```{r mnp-data, include = F}
mnp <- data.frame(
  cts.inst.type = c("c5.4xlarge", "c5.4xlarge", "c5.4xlarge", "c5.4xlarge", "c5.4xlarge"),
  cts.inst.num = c(1, 2, 3, 4, 5),
  rt1.inst.type = c("c5.2xlarge", "c5.2xlarge", "c5.2xlarge", "c5.2xlarge", "c5.2xlarge"),
  rt1.inst.num = c(2, 2, 2, 2, 3),
  tps.peak = c(506, 804, 1140, 1510, 1996),
  cost.mtx = c(1.67, 0.99, 0.66, 0.47, 0.35))
```

```{r mnp-data-table, echo = F}
kable(mnp, caption = "Multi-node pricing and TPS data",
  col.names = c("CTS inst. type", "CTS inst. number",
    "RT1 inst. type", "RT1 inst. number", "TPS peak", "1M TX cost"),
  align = "crcrrr")
```

### Performance of number of instances


```{r mnp-cts.inst.num-tps.peak, fig.cap = "TPS peak per CTS instance number"}
ggplot(mnp, aes(cts.inst.num, tps.peak,
    color = factor(cts.inst.num), fill = factor(cts.inst.num))) +
  geom_bar(stat = "identity", alpha = 0.2) +
  xlab("CTS instance number") + ylab("TPS peak")
```

### Cost of 1M transactions of number of instances


```{r mnp-cts.inst.num-cost.mtx, fig.cap = "Cost of 1M TX per CTS instance number"}
ggplot(mnp, aes(cts.inst.num, cost.mtx,
    color = factor(cts.inst.num), fill = factor(cts.inst.num))) +
  geom_bar(stat = "identity", alpha = 0.2) +
  xlab("CTS instance number") + ylab("Cost of 1M TX")
```

### Cost efficiency of number of instances


```{r mnp-cost.mtx-tps.peak, fig.cap = "TPS peak vs cost of 1M TX per CTS instance number"}
ggplot(mnp, aes(cost.mtx, tps.peak)) +
  geom_point(aes(size = tps.peak, color = tps.peak)) +
  scale_size_continuous(range = c(5, 12)) +
  geom_text(aes(x = cost.mtx, y = tps.peak,
    label = sprintf("(%d CTS, %d RT1)", cts.inst.num, rt1.inst.num)),
    hjust = "left", nudge_x = 0.1, size = 4) +
  xlim(0.3, 2.2) + xlab("Cost of 1M TX") + ylab("TPS peak")
```

# Conclusions

1. The most performant instsance type is c6g and the least peforman instance type is c5a
1. The most expensive instance type is c5n and the least expensive instance type is c6g
1. The highest cost efficiency is demonstrated by the c6g instance type, while the c5a
   instance has the smallest cost efficiency
1. Performance linearly increases with the number of instances
1. Cost of 1M TX decreases almost exponentially with the number of instances
1. Cost efficiency increases exponentially with the number of instances, so the most
   cost effective configuration is 5 CTS nodes and 3 RT1 nodes
