---
title: A minimal Rmd example with ggplot2
author: Volodymyr Prokopyuk
date: `r Sys.Date()`
abstract: |
  Markdown and knitr along with ggplot2 make literate programming in R very convenient
keywords: [markdown, knitr, literate-programming, ggplot2]
lang: en
css:
  - style/style.css
header-includes:
  - <link rel="shortcut icon" href="image/html5-favicon.png">
---

```{r document-config, include = F}
opts_chunk$set(error = F, dev = "svg")
```
<!--
```{r exports-data, include = T}
raw_export <- "year,product,export,percentage
2006,copper,4335009500,81
2006,others,1016726518,19
2007,copper,9005361914,86
2007,others,1523085299,14
2008,copper,6907056354,80
2008,others,1762684216,20
2009,copper,10529811075,81
2009,others,2464094241,19
2010,copper,14828284450,85
2010,others,2543015596,15
2011,copper,15291679086,82
2011,others,3447972354,18
2012,copper,14630686732,80
2012,others,3583968218,20
2013,copper,15244038840,79
2013,others,4051281128,21
2014,copper,14703374241,78
2014,others,4251484600,22
2015,copper,13155922363,78
2015,others,3667286912,22
"
export <- read_csv(raw_export)

# export <- export |>
#   mutate(product = factor(
#     product,
#     levels = c("copper", "others"),
#     labels = c("Copper", "Other goods")))

export |> mutate(product = factor(
  product,
  levels = c("copper", "others"),
  labels = c("Copper", "Other goods"))) -> export
```
-->

<!--
```{r exports-line-plot, fig.cap = "Exports to China", fig.dim = c(9, 6)}
ggplot(export, aes(x = year, y = export, color = product)) +
  geom_line(size = 1.5) + geom_point(size = 3) +
  scale_x_continuous(breaks = min(export$year):max(export$year)) +
  labs(title = "Exports to China, USD", subtitle = "Source: Observatory of economics") +
  labs(x = "Year", y = "USD") +
  theme(
    legend.title = element_blank(),
legend.position = "bottom",
    legend.direction = "horizontal") +
  scale_color_manual(values = c("#AA3939", "#7A9F35", "#226666"))
```
-->

<!--
```{r exports-area-plot, fig.cap = "Exports to China", fig.dim = c(9, 6)}
ggplot(export, aes(x = year, y = export, fill = product)) +
  geom_area(alpha = 0.7) +
  scale_x_continuous(breaks = min(export$year):max(export$year)) +
  labs(title = "Exports to China, USD", subtitle = "Source: Observatory of economics") +
  labs(x = "Year", y = "USD") +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.title = element_blank()) +
  scale_fill_manual(values = c("#E1B378","#5F9EA0"))
```
-->

<!--
```{r example-plot}
# ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point()
# ggplot(data = NULL, aes(x = mtcars$wt, y = mtcars$mpg)) + geom_point()
# ggplot(pressure, aes(x = temperature, y = pressure)) + geom_line() + geom_point()
# d <- tibble(
#   time = c(1, 2, 3, 4, 5, 7),
#   demand = c(8.3, 10.3, 19, 16, 15.6, 19.8),
#   name = c("a", "b", "c", "d", "e", "f"))
# ggplot(d, aes(x = time, y = demand)) + geom_col()
# ggplot(d, aes(x = factor(time), y = demand)) + geom_col()
# ggplot(d, aes(x = time, y = demand)) + geom_bar(stat = "identity")
# ggplot(mtcars, aes(x = cyl)) + geom_bar()
# ggplot(mtcars, aes(x = factor(cyl))) + geom_bar()
# ggplot(mtcars, aes(x = mpg)) + geom_histogram(binwidth = 2)
# ggplot(mtcars, aes(x = mpg)) + geom_histogram(bins = 10)
# ggplot(ToothGrowth, aes(x = supp, y = len)) + geom_boxplot()
# ggplot(ToothGrowth, aes(x = interaction(supp, dose), y = len)) + geom_boxplot()
# ggplot(tibble(x = -4:4), aes(x = x)) +
#   stat_function(fun = \(x) x^3 - 5 * x, geom = "line", color = "red") +
# ggplot(tibble(x = 0:20), aes(x = x)) +
#   stat_function(fun = \(x) 1 / (1 + exp(-x + 10)), geom = "line", color = "green")
# options(digits.secs = 6)
# a = Sys.time()
# Sys.sleep(0.001)
# b = Sys.time()
# b - a
```
-->

## Physical characteristics

The Moon is a very slightly scalene ellipsoid due to tidal stretching, with its long
axis displaced 30° from facing the Earth, due to gravitational anomalies from impact
basins. Its shape is more elongated than current tidal forces can account for. This
'fossil bulge' indicates that the Moon solidified when it orbited at half its current
distance to the Earth, and that it is now too cold for its shape to adjust to its
orbit.[56]

### Internal structure

The Moon is a differentiated body that was initially in hydrostatic equilibrium but has
since departed from this condition.[58] It has a geochemically distinct crust, mantle,
and core. The Moon has a solid iron-rich inner core with a radius possibly as small as
240 kilometres (150 mi) and a fluid outer core primarily made of liquid iron with a
radius of roughly 300 kilometres (190 mi). Around the core is a partially molten
boundary layer with a radius of about 500 kilometres (310 mi).[59][60] This structure is
thought to have developed through the fractional crystallization of a global magma ocean
shortly after the Moon's formation 4.5 billion years ago.[61]

Crystallization of this magma ocean would have created a mafic mantle from the
precipitation and sinking of the minerals olivine, clinopyroxene, and orthopyroxene;
after about three-quarters of the magma ocean had crystallised, lower-density
plagioclase minerals could form and float into a crust atop.[62] The final liquids to
crystallise would have been initially sandwiched between the crust and mantle, with a
high abundance of incompatible and heat-producing elements.[1] Consistent with this
perspective, geochemical mapping made from orbit suggests a crust of mostly
anorthosite.[14] The Moon rock samples of the flood lavas that erupted onto the surface
from partial melting in the mantle confirm the mafic mantle composition, which is more
iron-rich than that of Earth.[1] The crust is on average about 50 kilometres (31 mi)
thick.[1]

The Moon is the second-densest satellite in the Solar System, after Io.[63] However, the
inner core of the Moon is small, with a radius of about 350 kilometres (220 mi) or
less,[1] around 20% of the radius of the Moon. Its composition is not well understood,
but is probably metallic iron alloyed with a small amount of sulfur and nickel; analyses
of the Moon's time-variable rotation suggest that it is at least partly molten.[64] The
pressure at the lunar core is estimated to be 5 GPa.[65]

#### Magnetic field

The Moon has an external magnetic field of generally less than 0.2 nanoteslas,[66] or
less than one hundred thousandth that of Earth. The Moon does not currently have a
global dipolar magnetic field and only has crustal magnetization likely acquired early
in its history when a dynamo was still operating.[67][68] However, early in its history,
4 billion years ago, its magnetic field strength was likely close to that of Earth
today.[66] This early dynamo field apparently expired by about one billion years ago,
after the lunar core had completely crystallized.[66] Theoretically, some of the remnant
magnetization may originate from transient magnetic fields generated during large
impacts through the expansion of plasma clouds. These clouds are generated during large
impacts in an ambient magnetic field. This is supported by the location of the largest
crustal magnetizations situated near the antipodes of the giant impact basins.[69]

#### Surface geology

Main articles: Topography of the Moon, Geology of the Moon, Moon rock, and List of lunar
features Geological features of the Moon (near side / north pole at left, far side /
south pole at right) Topography of the Moon measured from the Lunar Orbiter Laser
Altimeter on the mission Lunar Reconnaissance Orbiter, referenced to a sphere of radius
1737.4 km Topography of the Moon

The topography of the Moon has been measured with laser altimetry and stereo image
analysis.[70] Its most extensive topographic feature is the giant far-side South
Pole–Aitken basin, some 2,240 km (1,390 mi) in diameter, the largest crater on the Moon
and the second-largest confirmed impact crater in the Solar System.[71][72] At 13 km
(8.1 mi) deep, its floor is the lowest point on the surface of the Moon.[71][73] The
highest elevations of the Moon's surface are located directly to the northeast, which
might have been thickened by the oblique formation impact of the South Pole–Aitken
basin.[74] Other large impact basins such as Imbrium, Serenitatis, Crisium, Smythii, and
Orientale possess regionally low elevations and elevated rims.[71] The far side of the
lunar surface is on average about 1.9 km (1.2 mi) higher than that of the near side.[1]

The discovery of fault scarp cliffs suggest that the Moon has shrunk by about 90 metres
(300 ft) within the past billion years.[75] Similar shrinkage features exist on
Mercury. Mare Frigoris, a basin near the north pole long assumed to be geologically
dead, has cracked and shifted. Since the Moon doesn't have tectonic plates, its tectonic
activity is slow and cracks develop as it loses heat.[76]

### Volcanic features

The dark and relatively featureless lunar plains, clearly seen with the naked eye, are
called maria (Latin for "seas"; singular mare), as they were once believed to be filled
with water;[77] they are now known to be vast solidified pools of ancient basaltic
lava. Although similar to terrestrial basalts, lunar basalts have more iron and no
minerals altered by water.[78] The majority of these lava deposits erupted or flowed
into the depressions associated with impact basins. Several geologic provinces
containing shield volcanoes and volcanic domes are found within the near side
"maria".[79]
