---
title: Luis. Presupuesto de la cubierta
author: Vlad
date: `r Sys.Date()`
abstract: >
  Presupuesto de la cubierta y del mirador
keywords: [techos, paredes, aislamiento, alisar]
lang: es
css:
  - style/style.css
header-includes:
  - <link rel="shortcut icon" href="image/html5-favicon.png">
---

```{r document-config, include = F}
opts_chunk$set(error = F, echo = F)
```

```{r datos-cubierta}
cubierta <- tribble(
  ~concepto, ~cantidad, ~precio,
  "Andamio", 160, 18,
  "Línea de vida", 1, 250,
  "Levantado de pizarra", 163, 10,
  # "Levantado de pizarra", 139, 10,
  "Enfoscado monocapa", 24, 50,
  # "Enfoscado monocapa", 6, 50,
  "Canalones y bajantes", 44, 35,
  # "Canalones y bajantes", 47, 35,
  "Baberos", 30, 25,
  # "Baberos", 19, 25,
  "Rastrel primario", 139, 12,
  "Telas asfálticas", 139, 25,
  "Rastrel secundario", 139, 15,
  "Teja plana", 139, 45,
  "Enfoscado chimenea", 1, 250,
  "Premarco de ventana", 1, 400,
  "Chapas perimetrales", 55, 18,
  "Tejas perimetrales", 55, 18,
  # Mirador
  "Demolición de tablero", 12, 15,
  "Desmontaje de ventana", 1, 50,
  "Picado de faldón", 10, 15,
  "Picado de muro", 6, 15,
  "Levantado de pavimento", 6, 15,
  "Picado de mortero", 10, 20,
  "Desmontaje de canalón", 1, 30,
  "Camara inferior", 29.5, 38,
  "Aislamiento", 30, 8,
  "Ladrillo cara vista", 21.7, 75,
  "Cargadero de sardinel", 3, 40,
  "Sardinel de ladrillo", 2.5, 75,
  "Camara superior", 17.5, 45,
  "Tubo de ventilación", 1, 30,
  "Vierteaguas", 3, 50,
  "Viga 5.5", 2, 350,
  "Viga 4.5", 1, 300,
  "Pilares 3.0", 2, 200,
  "Chapón", 2, 170,
  "Chapa colaborante", 24.8, 85,
  "Fibrocemento en aleros", 15.5, 35,
  "Monocapa en aleros", 15.5, 32)
cubierta |> mutate (valor = cantidad * precio) |>
  gt(rowname_col = "concepto") |>
  tab_header(title = "Presupuesto de la cubierta y del mirador") |>
  grand_summary_rows(columns = "valor", fns = list(total = "sum"))

# sum(cubierta$valor)
